### 2017-12-11 Draft2/Indexing.Growth.tex
Painful pondering

Clearly, $\FuncDecompose$ runs in constant time. Despite that, it appears to be quite expensive: normally, logarithms and exponentiation require use of floating-point instructions, which are slower than integer-based instructions. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.

To see this, first note that $\VarIndexBuffer$ becomes $\max(\left\lceil \log_2 (i + 1) \right\rceil - \VarLogInitCapacity, 0)$. There is a constant time algorithm to compute $\left\lceil \log_2 k \right\rceil$ without using floating-point for any $k \in \mathbb{N}$, which I will not discuss because it involves concepts outside the scope of this paper \cite{ceillog2}. Equipped with such an algorithm, it is easy to see that the whole expression can be computed without use of floating-point.

In calculating $\VarIndexElement$, the only potential use of floating-point instructions comes from $\VarGrowthFactor^{\VarIndexBuffer - 1}$. When $\VarGrowthFactor = 2$, however, this can be calculated with a simple bit shift. Thus, it is not necessary to use floating-point instructions to calculate either $\VarIndexBuffer$ or $\VarIndexElement$.

>>>>

At first glance, $\FuncDecompose$ appears to be expensive: logarithms and exponentiation normally imply use of floating-point instructions, which tend to be slower than integer-based ones. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.

To see this, note that $\VarIndexBuffer$ becomes $\max(\left\lceil \log_2 (i + 1) \right\rceil - \VarLogInitCapacity, 0)$. Using the concept of de Bruijn sequences, one can index the leftmost $1$ in the binary representation of $i$ in constant time \cite{ceillog2}. (For example, if $i = 9 = 1001\text{ in binary}$, the index of the leftmost $1$ starting from the right is $3$.) This index is the same as $\left\lfloor \log_2 i \right\rfloor$. Using the fact that $\left\lceil \log_2 (i + 1) \right\rceil = \left\lfloor \log_2 i \right\rfloor + 1$ (as there cannot be natural numbers between $\log_2 i$ and $\log_2 (i + 1)$), one can easily find $\VarIndexBuffer$. Finally, calculating $\VarGrowthFactor^{\VarIndexBuffer - 1}$ becomes a simple bit shift when $\VarGrowthFactor = 2$.
